---
title: "Applied Regression Analysis Week 2"
author: "Pashakhin Sergey"
date: "Saturday, April 04, 2015"
output: html_document
---

```{r}
setwd("~/applregress")
data <- read.csv("week2-HW-data.csv")
library(ggplot2)

```


## Exercise One:

Generate scatter diagrams for each of the following variable pairs:
SBP (Y) vs. QUET (X)
SBP (Y) vs. SMK (X)
QUET (Y) vs. AGE (X)
SBP (Y) vs. AGE (X)

```{r, echo=FALSE}
qplot(QUET, SBP, data=data)
qplot(SMK, SBP, data=data)
qplot(AGE, QUET, data=data)
qplot(AGE, SBP, data=data)
```

## Exercise Two:

Sketch a line
Using scatter diagrams #1, #3, and #4 that you generated above, use paper and pencil to roughly sketch a line that fits the data reasonably well. Use the homework forum to share your sketches and comment on the relationships described.

```{r, echo=FALSE}
qplot(QUET, SBP, data=data) + geom_smooth(method=lm)
qplot(AGE, QUET, data=data) + geom_smooth(method=lm)
qplot(AGE, SBP, data=data) + geom_smooth(method=lm)
```

## Exercise Three:

Comparing Blood Pressure with Smoking History

1. Determine the least-squares estimates of slope β1 and intercept β0 for the straight-line regression of SBP (Y) on SMK (X).

2. Compare the value of β^0 with the mean SBP for nonsmokers. Compare the value of β^0+β^1 with the mean SBP for smokers. Use the homework forum to explain and share the results of these comparisons.

3. Test the hypothesis that the true slope β1 is 0.

4. Is the test in question (3) equivalent to the usual two-sample t test for the equality of two population means assuming equal but unknown variances? Demonstrate your answer numerically and share your findings in the homework forum.

```{r}
## 1
ex31 <- lm(SBP ~ SMK, data=data)
summary(ex31)
```

```{r}
## 2
nonsmokers <- subset(data, SMK == 0)
mean(nonsmokers$SBP)
summary(ex31)$coefficients[1]

summary(ex31)$coefficients[1] + summary(ex31)$coefficients[2]
smokers <- subset(data, SMK == 1)
mean(smokers$SBP)
```

```{r}
## 3
t.test(SBP ~ SMK, data=data)
## 4
t.test(SBP ~ SMK, data=data, var.equal=T)
```

## Exercise Four:

Comparing Blood Pressure with Body Size

1. Determine the least-squares estimates of slope and intercept for the straight-line regression of SBP (Y) on QUET (X).

2. Sketch the estimated regression line on the scatter diagram involving SBP and QUET.

3. Test the hypothesis of zero slope.

4. Find a 95% confidence interval for μy|x¯.

5. Calculate 95% prediction bands.

6. Based on the above, would you conclude that blood pressure increases as body size increases? Share your explanation and reasoning in the homework forum.

7. Are any of the assumptions for straight-line regression clearly not satisfied in this example? Explain in the homework forum.

```{r}
## 1
ex32 <- lm(SBP ~ QUET, data=data)
summary(ex32)

## 2 
qplot(QUET, SBP, data=data) + geom_smooth(method=lm)

## 3
summary(ex32)  # The part I am confused with

## 4
newdata <- data.frame(QUET=data$QUET)
summary(predict(ex32, newdata=newdata, interval="confidence"))

## 5
summary(predict(ex32, newdata=newdata, interval="prediction"))
```

### 6
Yes.

### 7
We could plot residuals and do diagnostic plots but it is beyond this week's materials. So, I would just say: no.
